---
title: "Stan Match Outcome Model Notebook"
output: html_notebook
---

# Load packages 

```{r }
library(dplyr)
library(ggplot2)
library(rstan)
```

# Load data
```{r }
data <- read.csv(file = file.path('data','match_data.csv'))
data %>% tail(3)
```

```{r }
# Select the right columns
data <- data %>%
  select(season,del_position,del_defense,del_offense,del_winning,home_win) %>%
  na.omit()
```

```{r }
# split into a test and training data set
train_data <- data %>% filter(season < 2019)
test_data <- data %>% filter(season == 2019)
```

# Comile the stancode into an executable

```{r }
model <- stan_model('match_outcome_model.stan')
```

# Pass data and configuration parameters and sample

```{r }
# Make data into a list for stan
data_for_stan <- list()
data_for_stan$N <- nrow(train_data)
data_for_stan$C <- ncol(train_data) -2
data_for_stan$X_train <- train_data %>% select(-season,-home_win) %>% as.matrix()
data_for_stan$match_outcome <- train_data$home_win
data_for_stan$N_test <- nrow(test_data)
data_for_stan$X_test <- test_data %>% select(-season,-home_win) %>% as.matrix()
```

```{r }
# Sample from the posterior
samples <- sampling(object = model,
                    data = data_for_stan,
                    chains = 2,
                    iter = 2000,
                    seed = 1234)

```

# Inspect the results

```{r }
results_jpt <- extract(samples,par = c('intercept','coeffs')) %>% as.data.frame()
results_jpt
```

```{r }
results_summary <- summary(samples,par = c('intercept','coeffs'))$summary %>% as.data.frame()
results_summary
```

# Check divergences

```{r }
get_sampler_params(samples,inc_warmup = FALSE)[[1]][,'divergent__'] %>% sum()
get_sampler_params(samples,inc_warmup = FALSE)[[2]][,'divergent__'] %>% sum()
```

# Check pairs plot
```{r }
pairs(samples,pars = c('intercept','coeffs'))
```

# Interrogate results

```{r }
# Intercept
extract(samples,pars = 'intercept') %>%
  as.data.frame() %>%
  ggplot() +
  geom_histogram(aes(x = intercept),fill = 'transparent',color = 'black') + 
  xlab('Intercept value') + 
  ylab('Frequency')

```

```{r }
# Plot the coefficients
summary(samples,pars = 'coeffs')$summary %>%
  as.data.frame() %>%
  mutate(term = data %>% select(-home_win,-season) %>% names()) %>%
  ggplot() + 
  geom_segment(aes(x = `2.5%`,xend=`97.5%`,y = reorder(term,mean),yend = reorder(term,mean)))+
  geom_point(aes(x = mean, y = reorder(term,mean), color = 'red'))+
  geom_vline(aes(xintercept = 0),linetype = 'dashed')+
  xlab('Effect size')+
  ylab('')

```

# Test set performance

```{r }
summary(samples,pars = 'y_pred_test')$summary %>% head()
```

```{r }
prediction_summary <- summary(samples,pars = 'y_pred_test')$summary %>%
  as.data.frame() %>%
  mutate(actual_outcome = test_data$home_win) %>%
  arrange(desc(`50%`)) %>%
  mutate(game_id = 1:n())

```

```{r }
prediction_summary %>%
  ggplot() + 
  geom_segment(aes(y = `2.5%`,
                   yend = `97.5%`,
                   x = game_id,
                   xend = game_id,
                   color = factor(actual_outcome)),alpha = 0.3)+
  geom_point(aes(x = game_id, y = `50%`,color = factor(actual_outcome)))+
  coord_cartesian(xlim = c(0,nrow(prediction_summary)),ylim = c(0,1)) + 
  xlab('game_id')+
  ylab('Predicted prbability')

```

## Calcualte the range of out of sample accuracy

```{r }
test_pred_matrix <- extract(samples)$y_pred_test
test_pred_matrix[test_pred_matrix>0.5] = 1
test_pred_matrix[test_pred_matrix<=0.5] = 0 
actual_outcome = test_data$home_win

calc_accuracy <- function(est){
  return(mean(est == actual_outcome))
}
accuracy_range <- apply(test_pred_matrix,1,calc_accuracy)
hist(accuracy_range)

```

# Exercises:

## 1. Manually calculate the test set probabilities

```{r }
# Here we need to run a simulation where for every model instance (sample), we need to predict every y (match attendance). We will use matrix algebra to vectorise this.

inv_logit = function(x){
  return( exp(x)/(1+exp(x)))
}

# Measure the number of data points and number of samples
npoints = data_for_stan$N_test
niterations = length(extract(ep4_samples)$intercept)
ncoeffs = data_for_stan$C

#Perform theta * x. Result = n samples x n data points
theta_x_X = extract(ep4_samples)$coeffs %*% t(as.matrix(data_for_stan$X_test[1:npoints,])) 

#Create  n samples x n data points sized intercept
I = matrix(extract(ep4_samples)$intercept,ncol = 1)[,rep(1,each= npoints)]

# Add together to create the matrix of predicted y's
y_pred = inv_logit(theta_x_X + I)

# Summarise each point
prediction_summary <- apply(y_pred,2,function(x) quantile(x,probs = c(0.025,0.5,0.975))) %>% 
  t() %>% 
  as.data.frame() %>%
  mutate(actual_outcome = test_data$match_outcome) %>%
  arrange(desc(`50%`)) %>%
  mutate(game_id = 1:n())


# Plot
ggplot(data = prediction_summary) + 
  geom_segment(aes(y = `2.5%`,
                   yend = `97.5%`,
                   x = game_id, 
                   xend = game_id,
                   color = factor(actual_outcome) ),alpha = 0.3)+
  geom_point(aes(x = game_id, y = `50%`,color = factor(actual_outcome)))+
  coord_cartesian(xlim = c(0,npoints),ylim = c(0,1))+
  xlab("Actual")+
  ylab("Predicted")

```


## 2. Calculate the range of ROC AUCs of the model.


```{r }
library(pROC)
oos_pred_matrix <- extract(ep4_samples)$y_pred_test
oos_pred_matrix[oos_pred_matrix<0.5] = 0
oos_pred_matrix[oos_pred_matrix>=0.5] = 1
actual_outcome = test_data$home_win

calc_auc <- function(est){
  return(auc(actual_outcome,est))
}
auc_range <- apply(oos_pred_matrix,1,calc_auc)
hist(auc_range)

```
















